### 2017-09-12 pages/math/linear-algebra/includes/matrix-multi02.tex
Update

\begin{bmatrix}
  \color{red}a & \color{red}b & \color{red}c \\
  \color{green}d & \color{green}e & \color{green}f \\
  \color{purple}g & \color{purple}h & \color{purple}i \\
\end{bmatrix}
\begin{bmatrix}
  \color{blue}r & \color{orange}s & \color{yellow}t \\
  \color{blue}u & \color{orange}v & \color{yellow}w \\
  \color{blue}x & \color{orange}y & \color{yellow}z \\
\end{bmatrix} =
\begin{bmatrix}
  (\color{red}a  \color{blue}r + \color{red}b  \color{blue}u + \color{red}c\color{blue}  x) &
  (\color{red}a\color{orange}s + \color{red}b\color{orange}v + \color{red}c\color{orange}y) &
  (\color{red}a\color{yellow}t + \color{red}b\color{yellow}w + \color{red}c\color{yellow}z) &
  \\
  (\color{green}d  \color{blue}r + \color{green}e  \color{blue}u + \color{green}f\color{blue}  x) &
  (\color{green}d\color{orange}s + \color{green}e\color{orange}v + \color{green}f\color{orange}y) &
  (\color{green}d\color{yellow}t + \color{green}e\color{yellow}w + \color{green}f\color{yellow}z) &
  \\
  (\color{purple}d  \color{blue}r + \color{purple}e  \color{blue}u + \color{purple}f\color{blue}  x) &
  (\color{purple}d\color{orange}s + \color{purple}e\color{orange}v + \color{purple}f\color{orange}y) &
  (\color{purple}d\color{yellow}t + \color{purple}e\color{yellow}w + \color{purple}f\color{yellow}z) &
\end{bmatrix}