### 2017-09-25 pages/math/calc-3/quiz-generator/in/quiz03.tex

::: Math-252 Quiz 3
% 1, 2, 3, 4
Using \(P(-2,0,3),Q(1,2,4),R(-3,1,0)\),
\begin{enumerate}[a.]
  \item Find a vector orthogonal to the plane determined by \(P\), \(Q\) and \(R\).
  \item Find an equation of the plane passing through \(P\), \(Q\) and \(R\).
  \item Find the set of parametric equations for the line through $Q$ and parallel to $\vec a=\langle 4,-3,-2\rangle$.
  \item Find the distance from the point $(-4,-1,5)$ to the plane passing through \(P\), \(Q\) and \(R\).
\end{enumerate}
===
\begin{enumerate}[a.]
  \item \(\vec n=\vec{PQ}\times\vec{PR}=\langle -7,8,5\rangle\)
  \item \(-7x+8y+5z=29\)
  \item \(x=1+4t,y=2-3t,z=4-2t;t\in\mathbb{R}\)
  \item \(D=\frac{16}{\sqrt{138}}\)
\end{enumerate}