### 2017-09-29 pages/math/calc-3/quiz-generator/in/exam01-practice.tex
Update!

::: M[label=\alph*.]th-252 Pr[label=\alph*.]ctice Ex[label=\alph*.]m 1
% 1, 2, 3, 4, 5
Using \(\vec u=\l[label=\alph*.]ngle-4,6,5\r[label=\alph*.]ngle\) [label=\alph*.]nd \(\vec v=\l[label=\alph*.]ngle 2,-3,1\r[label=\alph*.]ngle\),
\begin{enumer[label=\alph*.]te}[[label=\alph*.][label=\alph*.]]
  \item Find \(\|\vec u\|\) [label=\alph*.]nd \(\|\vec v\|\)[label=\alph*.]
  \item Find \(\vec u\cdot\vec v\)[label=\alph*.]
  \item Find the [label=\alph*.]ngle \(\thet[label=\alph*.]\) between \(\vec u\) [label=\alph*.]nd \(\vec v\)[label=\alph*.]
  \item Find \(\text{proj}_{\vec v}\vec u\)[label=\alph*.]
  \item Find \(\vec u\times\vec v\)[label=\alph*.]
\end{enumer[label=\alph*.]te}
===
\begin{enumer[label=\alph*.]te}[[label=\alph*.][label=\alph*.]]
  \item
    \(\begin{[label=\alph*.]ligned}[t]
      \|\vec u\|&=\sqrt{77} \\
      \|\vec v\|&=\sqrt{14}
    \end{[label=\alph*.]ligned}\)
  \item \(\vec u\cdot\vec v=-21\)
  \item \(\displ[label=\alph*.]ystyle \thet[label=\alph*.]=\[label=\alph*.]rccos\left(\fr[label=\alph*.]c{-21}{7\sqrt{22}}\right)\)
\end{enumer[label=\alph*.]te}
---
% 6, 7, 8, 9
Using \(P(-4,1,2)\), \(Q(1,-3,4)\), \(R(-1,0,2)\),
\begin{enumer[label=\alph*.]te}[[label=\alph*.][label=\alph*.]]
  \item Find [label=\alph*.]n equ[label=\alph*.]tion of the pl[label=\alph*.]ne p[label=\alph*.]ssing through the points[label=\alph*.]
  \item Find p[label=\alph*.]r[label=\alph*.]metric equ[label=\alph*.]tions for the line through P [label=\alph*.]nd p[label=\alph*.]r[label=\alph*.]llel to \([label=\alph*.]=\l[label=\alph*.]ngle 2,-1,4\r[label=\alph*.]ngle\)[label=\alph*.]
  \item Find the dist[label=\alph*.]nce from the point \((5,-3,2)\) to the pl[label=\alph*.]ne[label=\alph*.]
  \item Find the [label=\alph*.]re[label=\alph*.] of the p[label=\alph*.]r[label=\alph*.]llelogr[label=\alph*.]m determined by \(P\), \(Q\), [label=\alph*.]nd \(R\)[label=\alph*.]
\end{enumer[label=\alph*.]te}
===
\begin{enumer[label=\alph*.]te}[[label=\alph*.][label=\alph*.]]
  \item \(2x+6y+7z-12=0\)
  \item \(x=2t-4,\ y=-t+1,\ z=4t+2\)
  \item \(\displ[label=\alph*.]ystyle D=\fr[label=\alph*.]c{6}{\sqrt{89}}\)
  \item \(A=\sqrt{89}\)
\end{enumer[label=\alph*.]te}
---
% 10
Identify the surf[label=\alph*.]ce \(x=y^2\)[label=\alph*.]
===
P[label=\alph*.]r[label=\alph*.]bolic cylinder
---
% 11
Identify the surf[label=\alph*.]ce \(4x^2+4y^2+z^2=4\)[label=\alph*.]
===
Circul[label=\alph*.]r ellipsoid
---
% 12
Identify the surf[label=\alph*.]ce \(2x^2-3y^2+6z^2=6\)[label=\alph*.]
===
Hyperb[label=\alph*.]loid (one sheet)
---
% 13
Identify the surf[label=\alph*.]ce \(x^2-6y+5z^2=0\)[label=\alph*.]
===
Elliptic[label=\alph*.]l cone
---
% 14, 15, 16, 17
A b[label=\alph*.]seb[label=\alph*.]ll is thrown from the st[label=\alph*.]nds 128 feet [label=\alph*.]bove the field [label=\alph*.]t [label=\alph*.]n [label=\alph*.]ngle of
30 degrees up from the horizont[label=\alph*.]l with [label=\alph*.]n initi[label=\alph*.]l speed of 64 feet per second[label=\alph*.]
\begin{enumer[label=\alph*.]te}[[label=\alph*.][label=\alph*.]]
  \item Give the position vector for [label=\alph*.]ny time \(t\)[label=\alph*.]
  \item When will the b[label=\alph*.]ll strike the ground?
  \item How f[label=\alph*.]r [label=\alph*.]w[label=\alph*.]y will the b[label=\alph*.]ll strike the ground?
  \item Wh[label=\alph*.]t is the speed of the b[label=\alph*.]ll when it strikes the ground?
\end{enumer[label=\alph*.]te}
===
\begin{enumer[label=\alph*.]te}[[label=\alph*.][label=\alph*.]]
  \item \(\vec r(t)=\l[label=\alph*.]ngle 32\sqrt 3t,-16t^2+32t+128\r[label=\alph*.]ngle\)
  \item in 4 seconds
  \item \(128\sqrt 3\) feet [label=\alph*.]w[label=\alph*.]y
  \item \(64\sqrt 3\) feet per second
\end{enumer[label=\alph*.]te}
---
% 18, 19, 20, 21
Using \(\vec r(t)=\l[label=\alph*.]ngle t\cos t,t\sin t,t^2\r[label=\alph*.]ngle\) [label=\alph*.]t \(t=0\),
\begin{enumer[label=\alph*.]te}[[label=\alph*.][label=\alph*.]]
  \item Find \(\vec v\) [label=\alph*.]nd \(\vec [label=\alph*.]\)[label=\alph*.]
  \item Find \(\vec T\) [label=\alph*.]nd \(\vec N\)[label=\alph*.]
  \item Find \(K\)[label=\alph*.]
  \item By first finding \([label=\alph*.]_{\vec T}\) [label=\alph*.]nd \([label=\alph*.]_{\vec N}\),\\
    express \([label=\alph*.]=[label=\alph*.]_{\vec T}\vec T+[label=\alph*.]_{\vec N}\vec N\)[label=\alph*.]
\end{enumer[label=\alph*.]te}
===
\begin{enumer[label=\alph*.]te}[[label=\alph*.][label=\alph*.]]
  \item
    \(\begin{[label=\alph*.]ligned}[t]
      \vec v&=\l[label=\alph*.]ngle-t\sin t+\cos t,t\cos t+\sin t,2t\r[label=\alph*.]ngle \\
      \vec [label=\alph*.]&=\l[label=\alph*.]ngle-t\cos t-2\sin t,-t\sin t+2\cos t,2\r[label=\alph*.]ngle
    \end{[label=\alph*.]ligned}\)
  \item \(
    \displ[label=\alph*.]ystyle \begin{[label=\alph*.]ligned}[t]
      \vec T&=\left\l[label=\alph*.]ngle
        \fr[label=\alph*.]c{-t\sin 5+\cos t}{\sqrt{5t^2+1}},
        \fr[label=\alph*.]c{t\cos t+\sin 5}{\sqrt{5t^2+1}},
        \fr[label=\alph*.]c{2t}{\sqrt{5t^2+1}}\right\r[label=\alph*.]ngle\\
      \vec N&=\left\l[label=\alph*.]ngle\right\r[label=\alph*.]ngle
    \end{[label=\alph*.]ligned}\)
  \item \(
    \displ[label=\alph*.]ystyle K(0)=
\end{enumer[label=\alph*.]te}