### 2017-10-21 pages/math/calc-3/quiz-generator/in/exam02-practice.tex
Add select quiz gen input files

::: Math-252 Exam 2
% 1
For \(f(x,y)=\sqrt{x^2-y^2}\) find the domain of \(f\) and describe the level
curves.
===
ANSWER
---
% 2
Find the limit \(\lim_{(x,y)\to(4,3)}\frac{\sqrt x-\sqrt{y+1}}{x-y-1},x\neq
y+1\).
===
ANSWER
---
% 3
Find \(f_{xy}\) for \(f(x,y)=\ln(xy+y^2)\).
===
ANSWER
---
% 4
If \(w=f(x,y)\), where \(x=r\cos\theta\) and \(y=r\sin\theta\), show that
\(f_x^2=f_y^2=(\frac{\delta w}{\delta r})^2+\frac{1}{r^2}(\frac{\delta
w}{\delta\theta})^2\).
===
ANSWER
---
% 5, 6, 7, 8
Using \(f(x,y)=\frac{x-y}{x+y}\) and \(P(2,-1)\),
\begin{enumerate}[label=\alph*.]
  \item Find the directional derivative of \(f\) in the direction of \(\vec
    v=\langle 4,-8\rangle\).
  \item Find the direction in which \(f\) increases most rapidly.
  \item Find the direction in which \(f\) decreases most rapidly.
  \item Find the maximum value of the directional derivative.
\end{enumerate}
===
\begin{enumerate}[label=\alph*.]
  \item ANSWER
  \item ANSWER
  \item ANSWER
  \item ANSWER
\end{enumerate}
---
% 9, 10
Using \(x^3-2xy+z^3+7y+6=0\) and \(P(1,4,-3)\),
\begin{enumerate}[label=\alph*.]
  \item Find an equation of the tangent plane at \(P\).
  \item Find equations of the normal line at \(P\).
\end{enumerate}
===
\begin{enumerate}[label=\alph*.]
  \item ANSWER
  \item ANSWER
\end{enumerate}
---
% 11
A flat metal plate lies on an xy-plane such that the temperature \(T\) at
\((x,y)\) is given by \(T=10(x^2+y^2)^2\), where \(T\) is in degrees and \(x\)
and \(y\) are in centimeters. Find the instantaneous rate of change of \(T\)
with respect to distance at \((1,2)\) in the direction of the x-axis.
===
ANSWER
---
% 12
The total resistance \(R\) of three resistances \(R_1\), \(R_2\) and \(R_3\)
connected in parallelis given by
\(\frac{1}{R}=\frac{1}{R_1}+\frac{1}{R_2}+\frac{1}{R_3}\). If measurements of
\(R_1\), \(R_2\) and \(R_3\) are 100, 200 and 400 ohms respectively, with a
maximum error of \(\pm 1\%\) in each measurement, estimate the maximum error in
the calculated value of \(R\).
===
ANSWER
---
% 13
Find the maximum and minimum values of \(f(x,y)=5+4x-2x^2+3y-y^2\) over the
triangular region with vertices \((0,0)\), \((2,0)\) and \((2,2)\).
===
ANSWER
---
% 14
Find the volume of the largest rectangular box that has three of its vertices on
the positive x, y and z-axes respectively, and a fourth vertex on the plane
\(3x+4y+2z=24\).
===
ANSWER
---
% 15
Reverse the order of integration of \(\int_1^e\int_0^{\ln x}y\ dy\ dx\) and
evaluate.