### 2018-05-09 assets/resume/resume.tex
Add files via upload

\documentclass[10pt, letterpaper]{article}

% -----------------------------------------------------------------------------
% Packages
% -----------------------------------------------------------------------------

% Fonts
\usepackage{fontspec}
%\setmainfont[SmallCapsFont = Questa SmallCaps]{Questa} % Main font
%\setmainfont{Questa Sans}

% Referencing
\usepackage{hyperref}
\usepackage[margin=1in]{geometry}
\usepackage{multicol}
\usepackage{censor}
\usepackage{titlesec}
\titleformat{\section}{\Large\scshape\raggedright}{}{0em}{}[\titlerule]
\titlespacing{\section}{0pt}{3pt}{3pt} % Spacing around sections
\titleformat{\subsection}{\scshape\raggedright}{}{0em}{}
\titlespacing{\subsection}{0pt}{3pt}{3pt}
\usepackage[parfill]{parskip}
\usepackage{tabu}
\pagestyle{empty}

% -----------------------------------------------------------------------------
% Document
% -----------------------------------------------------------------------------

\begin{document}

\StopCensoring{}

\textbf{\LARGE N\censor{ils} O\censor{lsson}}
\medskip

\begin{tabu}{l r}
  \begin{tabu}{r l}
    \textsc{Phone}    & \censor{(619) 292-1338} \\
    \textsc{Email}    & \censor{\href{mailto:nilso@enosis.net}{nilso@enosis.net}} \\
    \textsc{Address}  & \censor{2720 Gregory St, San Diego, CA 92104} \\
  \end{tabu} &
  \begin{tabu}{r l}
    \textsc{Website}  & \censor{\url{https://nilsso.github.io}} \\
    \textsc{Github}   & \censor{\url{https://github.com/nilsso}} \\
    \textsc{Linkedin} & \censor{\url{https://linkedin.com/nilsso}}
  \end{tabu}
\end{tabu}

\section{Objective}
((INSERT TAILORED OBJECTIVE MESSAGE HERE))
%To foster my software development and design sills within a professional,
%experienced, and cooperative team environment.
\section{Education}
\begin{multicols}{2}
  \begin{minipage}{0.5\textwidth}
    \subsection{San Diego Mesa College (2010--2018)}
    \begin{footnotesize}
      Degrees earned:
      \begin{itemize}
        \item
          Associate of Arts: Music Studies

        \item
          Associate of Arts: Mathematics
      \end{itemize}
    \end{footnotesize}
    \medskip

    \subsection{San Diego State University (2018--Present)}
    \begin{footnotesize}
      Degrees in progress (expected 2020):
      \begin{itemize}
        \item
          B.S. Applied Mathematics:\\
          \phantom{B.s. }Emphasis in Computational Science
      \end{itemize}
    \end{footnotesize}
  \end{minipage}
  \columnbreak%

  \begin{minipage}{0.5\textwidth}
    \subsection{Senior Tutor (2016--2018)}
    \begin{footnotesize}
      \begin{itemize}
        \item
          Senior music and computer science tutor at Mesa college
        \item
          Covered music theory, ear training, piano, voice, C/C++,
          Java, C\#, Microsoft VB, Excel and Access
      \end{itemize}
    \end{footnotesize}
    \medskip

    \subsection{San Diego City Robotics club (2013--2014)}
    \begin{footnotesize}
      \begin{itemize}
        \item
          Served as software team leader
        \item
          Participated in
          \href{http://www.robonation.org/competition/robosub}{RoboSub
          competition}, 2014 season
      \end{itemize}
    \end{footnotesize}
  \end{minipage}
\end{multicols}

\section{Computer skills}
\begin{multicols}{2}
  \begin{minipage}{0.5\textwidth}
    \subsection{Languages}
    \begin{footnotesize}
      \begin{itemize}
        \item
          C/C++, Java, Python, Lua, SQL and \LaTeX
      \end{itemize}
    \end{footnotesize}
    \medskip

    \subsection{Tools}
    \begin{footnotesize}
      \begin{itemize}
        \item
          Command line, Vim, Git, CMake, Gradle
        \item
          Major IDEs (Visual Studio, Eclipse, Netbeans, IntelliJ, etc.)
      \end{itemize}
    \end{footnotesize}
  \end{minipage}
  \columnbreak%

  \begin{minipage}{0.5\textwidth}
    \subsection{Other Technologies}
    \begin{footnotesize}
      \begin{itemize}
        \item
          APIs and Frameworks: Qt, OpenCV, SDL, NumPy
        \item
          Databases: MySQL, PostgresSQL, MariaDB
        \item
          Design patterns: MVC, singleton, command, OOP
        \item
          DevOps: AWS, Docker, Heroku
        \item
          Operating system: Linux, macOS and Windows
      \end{itemize}
    \end{footnotesize}
  \end{minipage}
\end{multicols}

\section{Projects}
\begin{footnotesize}
  \begin{itemize}
    \item
      \textsc{\href{https://nilsso.github.io}{Website}:}
      Houses my notes on a wide variety of subjects, including mathematics, music and programming

    \item
      \textsc{\href{https://github.com/nilsso/ACalculator}{Calcuator project}:}
      Final project for Data Structures in C++ class at Mesa
      College. Exemplifies several design patterns: commands, singletons, chain
      of responsibility
  \end{itemize}
\end{footnotesize}

\section{Other skills}
\begin{footnotesize}
  \begin{itemize}
    \item
      \textsc{Self starter:}
      Learned programming in 2010 via the game World of Warcraft

    \item
      \textsc{Musically inclined:}
      Spent several semesters singing in the Mesa College vocal ensemble; have
      gone on tours and performed solo in New York, Connecticut, Sacramento,
      Puebla Mexico, Poland and Czech Republic. Currently singing
      semi-professionally for a San Diego church
  \end{itemize}
\end{footnotesize}

\end{document}