### 2019-04-24 pages/math/semi-comb/project/paper.tex
Update

\documentclass{amsart}
\usepackage[T1]{fontenc}
\usepackage{mathpazo}
%\usepackage[margin=1in]{geometry}
\usepackage[inline]{enumitem}
%\usepackage{rotating}
%\usepackage{color}
%\usepackage{tabu}
%\usepackage{parskip}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{calc}
\usepackage{xfrac}
\usepackage{mathtools}
\usepackage{cancel}
\usepackage{tikz}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{calc}
%\usepackage{xcolor-solarized}
%\usepackage{minted}
%\usepackage{MnSymbol}
%\usepackage[libertine]{newtxmath}
%\newcommand*\circled[1]{\tikz[baseline= (char.base)]{
    %\node[shape=circle,draw,inner sep=2pt] (char) {#1};}}
%\usepackage{xypic}
%\usepackage[tightpage]{preview} % display
% Math operators
\DeclareMathOperator{\Z}{\mathbb{Z}}
\DeclareMathOperator{\Q}{\mathbb{Q}}
\DeclareMathOperator{\N}{\mathbb{N}}
\DeclareMathOperator{\R}{\mathbb{R}}
\DeclareMathOperator{\C}{\mathbb{C}}
\DeclareMathOperator{\F}{\mathbb{F}}
\DeclareMathOperator{\sge}{\mathsf{e}}
\DeclareMathOperator{\sgm}{\mathsf{m}}
\DeclareMathOperator{\sgG}{\mathsf{G}}
\DeclareMathOperator{\sgg}{\mathsf{g}}
\DeclareMathOperator{\sgF}{\mathsf{F}}
\DeclareMathOperator{\sgt}{\mathsf{t}}
\DeclareMathOperator{\sgL}{\mathsf{L}}
\DeclareMathOperator{\sgZ}{\mathsf{Z}}
\DeclareMathOperator{\sgAp}{Ap}
\DeclareMathOperator{\sgPF}{PF}
\DeclareMathOperator{\sgconv}{conv}
\DeclareMathOperator{\sgsupp}{supp}
\DeclareMathOperator{\calO}{\mathcal{O}}
\DeclareMathOperator{\Grho}{\prescript{}{\rho}G}
% Theorem environments
\theoremstyle{plain}
\newtheorem*{thm}{Theorem}
\newtheorem*{lemma}{Lemma}
\newtheorem*{claim}{Claim}
\newtheorem*{cor}{Corollary}
\newtheorem*{prop}{Proposition}
\newtheorem*{conj}{Conjecture}
\theoremstyle{remark}
\newtheorem*{example}{Example}
\theoremstyle{definition}
\newtheorem*{definition}{Definition}
% Paired math operators
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
% Math helper functions
\newcommand{\congr}[2]{{[#1]}_{#2}}
\newcommand{\set}[1]{\{#1\}}
\newcommand{\genset}[1]{{\langle{}#1\rangle{}}}
\newcommand{\abs}[1]{\lvert{}#1\rvert{}}
\newcommand{\norm}[1]{\lVert{}#1\rVert{}}
\newcommand{\bmx}[1]{\begin{bmatrix}#1\end{bmatrix}}
\newcommand{\pmx}[1]{\begin{pmatrix}#1\end{pmatrix}}
  \newcommand{\sbmx}[1]{{\left[\begin{smallmatrix}#1\end{smallmatrix}\right]}}
\newcommand{\smallxymatrix}[2]{\xymatrix@C-=#1cm@R-=#1cm{#2}}
\renewcommand\labelitemi{\small$\bullet$}
\begin{document}
%
%
%
\noindent
\begin{quote}
  (13) \emph{Flow polytopes.}
  A \emph{directed graph} is a collection of dots (called \emph{vertices}) and
  arrow between them (called \emph{edges}). A \emph{flow} on a directed graph is a
  way to label each edge of $G$ by an element of $\Z$ (or $\Z_n$) so that at each
  vertex $v$, the sum of the edges entering $v$ equals the sum of the edges
  leaving $v$ (i.e.\ the in \emph{flow} at each vertex equals the \emph{out
  flow}). The flows on a given directed graph can be counted using a certain
  family of polytopes, whose integer points each coincide with a flow on $G$.
  \\[1em]
  \noindent
  Source:
  \emph{Combinatorial reciprocity theorems} (M. Beck, R. Sanyal), Chapters 1 and 7.
\end{quote}

\hrulefill

% 1.2 Flows on Graphs notes

A graph $G=(V,E)$ is a collection of $V$ a set of vertices and $E$ a set of
edges between these vertices. Now, a graph doesn't have to have any vertices or
any edges (except that if it has edges there must be vertices to which those edges
connect).
Also note that the vertices of a graph can be elements of some arbitrary $\F^n$
space, but for this report we'll only concern ourselves with graphs in $\Z^2$
Euclidean space.
\begin{definition}[Component, connected component]
  A \emph{connected component} (or simply a \emph{component}) of a graph $G$ is
  a subgraph (a subset of the vertices and edges of $G$) in which any two
  vertices are either directed connect via an edge or are connected to
  one-another via a \emph{path} (a series of vertices and edges). A single
  vertex with no connecting edges is in itself a connected component.
\end{definition}
Any graph then is the disjoint union of connected components.
For example figure~\ref{fig:disjoint-union} can be called a single graph that is the disjoint
union of three distinct connected components.
\begin{figure}[h]
  \begin{tikzpicture}[every node/.style={draw,shape=circle,fill=white,inner sep=1}]
    \path
      (.5,.5) node (a) {}
      (0,0) node (b) {} (1,0) node (c) {} (1,1) node (d) {} (0,1) node (e) {}
      (2,1) node (f) {} (2,0) node (g) {} (3,0) node (h) {};
    \draw (a);
    \draw (b) -- (c) -- (d) -- (e) -- (b);
    \draw (f) -- (g) -- (h);
  \end{tikzpicture}
  \caption[Disjoin union]{}
  \label{fig:disjoint-union}
\end{figure}
\begin{lemma}
  A connected component of the graph $G$ is a maximal subgraph of $G$ in which
  any two nodes are connect by a path.
\end{lemma}
\begin{lemma}
  A graph $G$ is \emph{connected} if it has only one connected component.
\end{lemma}
\begin{definition}[Bridge]
  A \emph{bridge} is a single edge which serves as a connection between two connected
  components which if removed would increase the number of connected
  components of the graph by one.
\end{definition}
For example, removing the bridge in figure~\ref{fig:bridge} turns a single
connected component into two triangular connected components:
one single connected graph into a graph with two connected subgraphs.
\begin{figure}[h]
  \begin{tikzpicture}[every node/.style={draw,shape=circle,fill=white,inner sep=1}]
    \path
      (-.5,.5) node (a) {} ($({-(1+sqrt(3))/2},0)$) node (b) {} ($({-(1+sqrt(3))/2},1)$) node (c) {}
      ( .5,.5) node (d) {} ($({ (1+sqrt(3))/2},0)$) node (e) {} ($({ (1+sqrt(3))/2},1)$) node (f) {};
    \draw (a) -- (b) -- (c) -- (a);
    \draw (d) -- (e) -- (f) -- (d);
    \draw[dashed] (a) -- (d);
  \end{tikzpicture}
  \hspace{1cm}
  \begin{tikzpicture}[every node/.style={draw,shape=circle,fill=white,inner sep=1}]
    \path
      (-.5,.5) node (a) {} ($({-(1+sqrt(3))/2},0)$) node (b) {} ($({-(1+sqrt(3))/2},1)$) node (c) {}
      ( .5,.5) node (d) {} ($({ (1+sqrt(3))/2},0)$) node (e) {} ($({ (1+sqrt(3))/2},1)$) node (f) {};
    \draw (a) -- (b) -- (c) -- (a);
    \draw (d) -- (e) -- (f) -- (d);
  \end{tikzpicture}
  \caption[Bridges]{A graph with and without a bridge.}
  \label{fig:bridge}
\end{figure}
So far we've only looked at graph which are \emph{unoriented}.
\begin{definition}[Directed, oriented graph]
  A \emph{directed} or \emph{oriented graph} is a graph is a graph on which we
  define an \emph{orientation} via a subset $\rho$ of the edges $E$ of the
  graph. For each edge $e=v_i v_j\in E$ (connecting the vertices $v_i,v_j\in
  V$) with $i<j$ we direct:
  \[
    v_i\stackrel{e}{\leftarrow}v_j\text{ if }e\in\rho\text{, and }
    v_i\stackrel{e}{\rightarrow}v_j\text{ if }e\notin\rho.
  \]
\end{definition}
This definition may be confusing at first glance because $\rho$
may be missing some of the edges of $E$. What this definition attempts to succinctly
encapsulate is only the edges with their direction pointing from the vertex of higher
index to the vertex of lower index (which does certainly feel backwards),
with all the remaining edges not in $\rho$ directed in the opposite direction,
from lower to higher index.
Figure~\ref{fig:directed-graph} tries to illustrate this encapsulation for a graph
with five vertices and an orientation $\rho=\{14,23,24\}$.
\begin{figure}[h]
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \begin{tikzpicture}[every node/.style={draw,shape=circle,fill=white,inner sep=1}]
      \path
        (0,1) node (v_1)[label=left:$v_1$]{}
        (1,1) node (v_2)[label=right:$v_2$]{}
        (1,0) node (v_3)[label=right:$v_3$]{}
        (0,0) node (v_4)[label=left:$v_4$]{};
      \draw (v_4) -- (v_3) -- (v_2) -- (v_1) -- (v_4) -- (v_2);
    \end{tikzpicture}
    \caption{The undirected graph.}
  \end{subfigure}
  %\hspace{1cm}
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \begin{tikzpicture}[
        every node/.style={draw,shape=circle,fill=white,inner sep=1},
        decoration={markings,mark=at position 0.5 with {\arrow{>}}}
      ]
      \path
        (0,1) node (v_1)[label=left:$v_1$]{}
        (1,1) node (v_2)[label=right:$v_2$]{}
        (1,0) node (v_3)[label=right:$v_3$]{}
        (0,0) node (v_4)[label=left:$v_4$]{};
      \draw[postaction={decorate}] (v_4) -- (v_1);
      \draw[postaction={decorate}] (v_4) -- (v_2);
      \draw[postaction={decorate}] (v_3) -- (v_2);
      \draw (v_3) -- (v_4);
      \draw (v_1) -- (v_2);
    \end{tikzpicture}
    \caption{$\rho$ edges marked.}
  \end{subfigure}
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \begin{tikzpicture}[
        every node/.style={draw,shape=circle,fill=white,inner sep=1},
        decoration={markings,mark=at position 0.5 with {\arrow{>}}}
      ]
      \path
        (0,1) node (v_1)[label=left:$v_1$]{}
        (1,1) node (v_2)[label=right:$v_2$]{}
        (1,0) node (v_3)[label=right:$v_3$]{}
        (0,0) node (v_4)[label=left:$v_4$]{};
      \draw[postaction={decorate}] (v_4) -- (v_1);
      \draw[postaction={decorate}] (v_4) -- (v_2);
      \draw[postaction={decorate}] (v_3) -- (v_2);
      \draw[postaction={decorate}] (v_3) -- (v_4);
      \draw[postaction={decorate}] (v_1) -- (v_2);
    \end{tikzpicture}
    \caption{$E\setminus\rho$ edges marked.}
  \end{subfigure}
  \caption[A directed graph]{A directed graph.}
  \label{fig:directed-graph}
\end{figure}

As an aside, directed graphs are the subject of many algorithms in computer
science, often pertaining to finding optimal ways to traverse the paths of said
graphs. In some models each edge is considered to be of equal ``weight'' as any
other. That is if two edges originating from the same node both point to the
same node, there should be no real preference to traveling down one edge or the
other to get to the destination. On the other hand we might assign a weight to
each edge, presenting a cost-benefit analysis in choosing a path to take (as in
figure~\ref{fig:path-finding} where the edge with weight 5 now becomes most
expensive to traverse versus the one with weight 2).
\begin{figure}[h]
  \begin{tikzpicture}[
    dot/.style={draw,shape=circle,fill=white,inner sep=1},
    decoration={markings,mark=at position 0.5 with {\arrow{>}}}
  ]
    \path (0,0) node[dot] (a){} (2,0) node[dot] (b){};
    \draw[postaction=decorate] (a) to [out=45, in=135] node[above]{} (b);
    \draw[postaction=decorate] (a) to [out=315,in=225] node[below]{$\phantom 5$} (b);
  \end{tikzpicture}
  \hspace{1cm}
  \begin{tikzpicture}[
    dot/.style={draw,shape=circle,fill=white,inner sep=1},
    decoration={markings,mark=at position 0.5 with {\arrow{>}}}
  ]
    \path (0,0) node[dot] (a){} (2,0) node[dot] (b){};
    \draw[postaction=decorate] (a) to [out=45, in=135] node[above]{$2$} (b);
    \draw[postaction=decorate] (a) to [out=315,in=225] node[below]{$5$} (b);
  \end{tikzpicture}
  \caption{A directed graph with and without weights.}
  \label{fig:path-finding}
\end{figure}

Intuitively these weights are often simple $\Z$ integers.
But moving on, we now we introduce \emph{flow} values to edges; similar
values to weights, but living not within $\Z$ but instead within an Abelian
group $\Z_n=\sfrac\Z n$. (There may also be optimal path algorithms associated
with flow values, but for this paper our focus is elsewhere.)
\begin{definition}[$\Z_n$-flow]
  A \emph{$\Z_n$-flow} is a map $f:E\to\Z_n$ which to each $e$ edge of $E$
  assigns a flow ${f(e)\in\Z_n}$ such that there is ``conservation of flow'' at
  every $v\in V$ vertex, by which me mean the sum of the flows into a vertex $v$
  is equivalent to the sum of the flows out it:
  \[
    \sum_{\stackrel e \to v}f(e)=\sum_{v\stackrel e \to}f(e).
  \]
\end{definition}
And it's important to note equivalence here since $\sum f(e)\in\Z_n$.
\begin{definition}[$f$ support]
  The \emph{support} of $f$, denoted $\sgsupp(f)$, is the subset of edges $e$
  whose flows are non-zero:
  \[
    \sgsupp(f)=\{e\in E:f(e)\ne 0\}\subseteq E
  \] 
\end{definition}
If $\sgsupp(f)=E$ (all $e$ edges have $f(e)\ne0$), then $f$ is considered
``nowhere-zero''. For the rest of this report we'll be concerned with these
nowhere-zero $\Z_n$-flows, in particular with counting how many of them can be
made for any particular graph and Abelian group.
(So going forward, $G$ will always refer to a graph in $\Z^2$ with $V$ vertices
and $E$ edges, $\rho$ to an arbitrary (but fixed) orientation of $G$,
and $\Z_n$ an Abelian group to which $f$ a $\Z_n$-flow maps.)
\begin{definition}
  Define a counting function:
  \[
    \varphi_G(n)=\text{the number of $f$ nowhere-zero $\Z_n$-flows on $G$}
  \] 
\end{definition}
As it happens\ldots
\begin{prop}
  The flow-counting function $\varphi_G(n)$ is independent of the orientation
  $\rho$ of $G$.
\end{prop}
(Todo:? Describe the proof of this proposition.)
\begin{lemma}
  $G$ will not have any nowhere-zero flow if $G$ has a \emph{bridge}---an edge
  whose removal increases the number of connected components of $G$.
\end{lemma}

% 7.6 Graph Flows and Totally Cyclic Orientations notes

%
%
%
\end{document}